---
title: "Projeto MEI - Meta 1"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Dependencies

```{r Libraries}
#install.packages("scatterplot3d", dependencies = TRUE)
library(scatterplot3d)
#install.packages("rgl")
library(rgl)
options(rgl.printRglwidget = TRUE)
```

```{r Compute colormap function}
compute_colormap <- function(d) {
  zlim <- range(d)
  zlen <- zlim[2] - zlim[1] + 1
  colorlut <- terrain.colors(zlen)
  colorlut[d - zlim[1] + 1]
}
```

# Max Flow problem dataframe

The dataset has the following parameters:

-   n_vert = number of vertices
-   prob = arc probability (0 \<= p \<= 1)
-   range = maximum range of capacity
-   seed = random seed
-   Dinic_time = time in secs
-   EK_time = time in secs
-   MPM_time = time in secs

```{r Load data set}
df <- read.csv(file = "resultsEDAv3.csv", header=TRUE)
```

# Data visualization

```{r}
summary(df)
```

## Dinic

```{r}
plot(df$Dinic_time ~ df$n_vert, xlab = 'n_vert', ylab = 'Dinic_time',
     main = 'Dinic.cpp')
plot(df$Dinic_time ~ df$prob, xlab = 'prob', ylab = 'Dinic_time',
     main = 'Dinic.cpp')
plot(df$EK_time ~ df$n_vert, xlab = 'n_vert', ylab = 'EK_time',
     main = 'EK.cpp')
plot(df$EK_time ~ df$prob, xlab = 'prob', ylab = 'EK_time',
     main = 'EK.cpp')
plot(df$MPM_time ~ df$n_vert, xlab = 'n_vert', ylab = 'MPM_time',
     main = 'MPM.cpp')
plot(df$MPM_time ~ df$prob, xlab = 'prob', ylab = 'MPM_time',
     main = 'MPM.cpp')
```

```{r}
plot3d(x = df$n_vert, y = df$prob, z =df$Dinic_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'Dinic_time',
       main = 'Dinic.cpp') 
plot3d(x = df$n_vert, y = df$prob, z =df$EK_time,
        xlab = 'n_vert', ylab = 'prob', zlab = 'EK_time',
        main = 'EK.cpp') 
plot3d(x = df$n_vert, y = df$prob, z =df$MPM_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'MPM_time',
       main = 'MPM.cpp')
```

# Linear Regression

## Dinic

```{r Dinic regression using linear model}
lr.out = lm(Dinic_time ~ I(n_vert * prob) + n_vert, data = df) 
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$Dinic_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'Dinic.cpp',
       col=c("blue"))
col <- compute_colormap(pred)  
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```

```{r Dinic regression using quadratic model}
lr.out = lm(Dinic_time ~ I(n_vert * prob) + I(n_vert^2), data = df)  
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$Dinic_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'Dinic.cpp',
       col=c("blue"))
col <- compute_colormap(pred)  
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```

```{r Dinic regression using cubic model}
lr.out = lm(Dinic_time ~ I(n_vert * prob) + I(n_vert^3), data = df) 
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$Dinic_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'Dinic.cpp',
       col=c("blue"))
col <- compute_colormap(pred)
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```
The quadratic model $A + n\_vert^2$ or $n\_vert * prob + n\_vert^2$ is clearly the better.

## EK

```{r EK regression using linear model}
lr.out = lm(EK_time ~ I(n_vert * prob) + n_vert, data = df) 
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$EK_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'EK.cpp',
       col=c("blue"))
col <- compute_colormap(pred)  
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```

```{r EK regression using quadratic model}
lr.out = lm(EK_time ~ I(n_vert * prob) + I(n_vert^2), data = df)  
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$EK_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'EK.cpp',
       col=c("blue"))
col <- compute_colormap(pred)  
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```

```{r EK regression using cubic model}
lr.out = lm(EK_time ~ I(n_vert * prob) + I(n_vert^3), data = df) 
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$EK_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'EK.cpp',
       col=c("blue"))
col <- compute_colormap(pred)  
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```

## MPM

```{r MPM regression using linear model}
lr.out = lm(MPM_time ~ I(n_vert * prob) + n_vert, data = df) 
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$MPM_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'MPM.cpp',
       col=c("blue"))
col <- compute_colormap(pred)  
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```

```{r MPM regression using quadratic model}
lr.out = lm(MPM_time ~ I(n_vert * prob) + I(n_vert^2), data = df)  
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$MPM_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'MPM.cpp',
       col=c("blue"))
col <- compute_colormap(pred)  
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```

```{r MPM regression using cubic model}
lr.out = lm(MPM_time ~ I(n_vert * prob) + I(n_vert^3), data = df) 
summary(lr.out)
plot(lr.out)
n_gen <- seq(min(df$n_vert), max(df$n_vert), length.out = 100)
p_gen <- seq(min(df$prob), max(df$prob), length.out = 100)
pred <- outer(X=n_gen, Y=p_gen, FUN=function(x,y)predict(lr.out, data.frame(n_vert = x, prob = y), type = "response"))
plot3d(x = df$n_vert, y = df$prob, z = df$MPM_time,
       xlab = 'n_vert', ylab = 'prob', zlab = 'MPM.cpp',
       col=c("blue"))
col <- compute_colormap(pred)  
surface3d(x = n_gen, y = p_gen, z = pred, color = col, back = "lines")
```

# Notas

```{r}
plot(lm((df$Dinic_time)~df$A+df$n_vert)) 

plot(lm((df$Dinic_time)~df$A+I(df$n_vert^2)))

plot(lm((df$Dinic_time)~df$A+I(df$n_vert^3)))
```

```{r}
scatterplot3d(x = df$n_vert, y =df$prob, z =df$Dinic_time,
              xlab = 'n_vert', ylab = 'prob', zlab = 'cpu_time',
              main = 'Dinic.cpp')

scatterplot3d(x = df$n_vert, y =df$prob, z =df$EK_time,
              xlab = 'n_vert', ylab = 'prob', zlab = 'cpu_time',
              main = 'EK.cpp')

scatterplot3d(x = df$n_vert, y =df$prob, z =df$MPM_time,
              xlab = 'n_vert', ylab = 'prob', zlab = 'cpu_time',
              main = 'MPM.cpp')
```

```{r}
boxplot(df$Dinic_time~df$n_vert) 
boxplot(df$EK_time~df$n_vert) 
boxplot(df$MPM_time~df$n_vert) 
```
